### 2017-11-10 prezzy/TAMU2017.tex
post TAMU changes to Slides
This gets at Flo’s issue on uTv and my concerns about estimation.

}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Latent Factor Model}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{Latent Factor Model: The ``M'' in AME}

@@

Can account for both stochastic equivalence and homophily

>>>>

}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Latent Factor Model}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\frame{
  \frametitle{What Exactly Is Third-Order Dependency?}
  \framesubtitle{People Want to Know!}
\begin{itemize}
\item<1> If each node has individual characteristics, 
say $x_i$ and $x_j$ for two nodes.  Then,
$x_i \times x_j$ is the multiplicative product of their nodal characteristics and may be related to their relationship, 
$y_{ij}$.  
Let $x$ be an encoding of whether a node is democratic or not. 
Then, $x_i \times x_j \equiv 1 \iff$ 
\textit{both are democracies}.  This approach is often used in standard analyses as a variable often known as 
\textit{joint democracy}.  In the context of networks, a positive effect from the interaction of $x_i$ and $x_j$ might affect some dyadic variable, such as not being in a militarized dispute against one another---often denoted the democratic peace.  This positive effect is known in network terminology as \textit{homophily}.
Most studies just look at the correlation between aggregated variables, but the proposed approach can more directly estimate the degree of homophily.
\end{itemize}
}

\frame{
  \frametitle{What Exactly Is Third-Order Dependency?}
  \framesubtitle{People Want to Know!}
\begin{itemize}
\item<1> Homophily which stems from nodal characteristics results in certain types of network structures: 1) transitivity, 2) balance, and 3) clustering.  For example, homophily tends to induce a large number of \textit{triads}, since if $i$ is similar to $j$ and as a result has a link with $j$, for the same reason it will share a link with $k$ if $j$ and $k$ are similar, thus creating a triad.  Groups of triads tend to result in \textit{clusters}. This can result in clusters in which many elements in the cluster are not only related to one another but share similar characteristics.  This is known as \textit{stochastic equivalence}.

\item To capture this kind of stochastic equivalence---and other types of triadic dependence---we include a latent representation of nodal characteristics in the regression model.  This allows the network representation in a matrix to be dependent upon how similar these latent characteristics are. These latent nodal characteristics are represented by a type of multiplicative effects that are expressed as $\mathbf{u}_i^\intercal \mathbf{v}_i$.  This captures the third-order dependencies, i.e., the dependencies among triads.
\end{itemize}
}

\frame{
  \frametitle{Latent Factor Model: The ``M'' in AME}

@@

This can account for both stochastic equivalence and homophily, as well as clustering and balance.

\frame{
  \frametitle{How Does Estimation Work?}
  \framesubtitle{It Is NOT All in the Error Term!}
 \begin{itemize}
  \item sample $\bm\theta \; | \;  \bm\beta, \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
  \item sample $\bm\beta \; | \;  \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
  \item sample $\mathbf{a}, \mathbf{b} \; | \; \bm\beta, \mathbf{X}, \bm\theta, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
  \item sample $\Sigma_{ab} \; | \; \bm\beta, \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Inverse-Wishart)
  \item update $\rho$ using a Metropolis-Hastings step with proposal $p^{*} | p  \sim$ truncated normal$_{[-1,1]}(\rho, \sigma_{\epsilon}^{2})$
  \item sample $\sigma_{\epsilon}^{2} \; | \; \bm\beta, \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \text{ and } \rho$ (Inverse-Gamma)
  \item For each $k \in K$:
  \begin{itemize}
    \item Sample $\mathbf{U}_{[,k]} \; | \; \bm\beta, \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}_{[,-k]}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
    \item Sample $\mathbf{V}_{[,k]} \; | \; \bm\beta, \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}_{[,-k]}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
    \item Sample $\mathbf{D}_{[k,k]}  \; | \; \bm\beta, \mathbf{X}, \bm\theta,\mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)\footnote{Subsequent to estimation, $\mathbf{D}$ matrix is absorbed into the calculation for $\mathbf{V}$ as we iterate through $K$. }
  \end{itemize}
 \end{itemize}
